---
title: "PA1_template.Rmd"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#STEP1

#Loading and pre processing the dataset
```{r}
activity <- read.csv("activity.csv")

# Overview of the activity data set

head(activity)
dim(activity)
str(activity)

#Looking at datea set we find the we need to transform the date column into date format

library(lubridate)
activity$date<-ymd(activity$date)
length(unique(activity$date))

summary(activity)
```
#__________________________________________________________________________________________

#STEP2:Histogram of the total number of steps taken each day


```{r}
activity_total_steps <- with(activity, aggregate(steps, by = list(date), FUN = sum, na.rm = TRUE))
names(activity_total_steps) <- c("date", "steps")
hist(activity_total_steps$steps, main = "Histogram of the total number of steps taken each day", xlab = "Days", ylab="Count", col = "darkblue", ylim = c(0,20), breaks = seq(0,25000, by=2500))
```

#__________________________________________________________________________________________

#STEP3: Mean and median number of steps taken each day

```{r}
su <- tapply(activity$steps, activity$date, sum, na.rm=T)
mean_su <- round(mean(su))
median_su <- round(median(su))
print(c("Mean",mean_su))
print(c("Median",median_su))
```

#__________________________________________________________________________________________

#STEP4: Time series plot of the average number of steps taken

```{r}
meanStepsInterval <- tapply(activity$steps, activity$interval, mean, na.rm=T)
plot(meanStepsInterval ~ unique(activity$interval), type="l", xlab = "5-min interval", main="Time series plot of the average number of steps taken")
```

#___________________________________________________________________________________________

#STEP5: The 5-minute interval that, on average, contains the maximum number of steps

```{r}
meanStepsInterval[which.max(meanStepsInterval)]
```

#___________________________________________________________________________________________

#STEP6: Code to describe and show a strategy for imputing missing data

```{r}
activity2 <- activity
for (i in 1:nrow(activity)){
    if(is.na(activity$steps[i])){
        activity2$steps[i]<- meanStepsInterval[[as.character(activity[i, "interval"])]]
    }
}

```

#___________________________________________________________________________________________

#STEP7:Histogram of the total number of steps taken each day after missing values are imputed

```{r}
finnnn <- tapply(activity2$steps, activity2$date, sum, na.rm=T)
hist(finnnn, xlab = "Days", ylab = "Count" , col = "darkblue", main = "Histogram of the total number of steps taken after missing values are imputed")

# Mean and median after imputing missing values with their 5 minute interval average
mean_su2 <- round(mean(finnnn))
median_su2 <- round(median(finnnn))
print(c("New Mean",mean_su2))
print(c("New Median",median_su2))
```

#___________________________________________________________________________________________

#STEP8:Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r}
Sys.setenv(LANGUAGE = "en")
Sys.setlocale("LC_TIME", "English")
activity2$weekday <- c("weekday")
activity2[weekdays(as.Date(activity2[, 2])) %in% c("Saturday", "Sunday", "saturday", "sunday"), ][4] <- c("weekend")
activity2$weekday <- factor(activity2$weekday)

activity2_weekend <- subset(activity2, activity2$weekday == "weekend")
activity2_weekday <- subset(activity2, activity2$weekday == "weekday")
mean_activity2_weekday <- tapply(activity2_weekday$steps, activity2_weekday$interval, mean)
mean_activity2_weekend <- tapply(activity2_weekend$steps, activity2_weekend$interval, mean)
library(lattice)
df_weekday <- data.frame(interval = unique(activity2_weekday$interval), avg = as.numeric(mean_activity2_weekday), day = rep("weekday", length(mean_activity2_weekday)))
df_weekend <- data.frame(interval = unique(activity2_weekend$interval), avg = as.numeric(mean_activity2_weekend), day = rep("weekend", length(mean_activity2_weekend)))
df_final <- rbind(df_weekday, df_weekend)
xyplot(avg ~ interval | day, data = df_final, layout = c(1, 2), type = "l", ylab = "Average Number of Steps")
```

